### 2020-06-05 dataprep.tex
Corrected DBE description.
preparing for OLCF import

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Data Collection and Preprocessing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Collection and Preprocessing}
\label{section:dataprep}
Data on Titan GPU life times is constructed from two sources:
inventory runs and failure event \rev{log} records. Two types of
failure events were collected \rev{for this study}: Double Bit Error
(DBE) and Off the Bus (OTB). \rev{DBE is a memory error detected by
  parity bit checking, which can correct a single bit flip and detect
  but not correct a double bit flip. OTB is the loss of host CPU
  connection to the GPU. While other types of errors were recorded in
  log files, as is reported in \cite{7266836,tiwari15Understand}, OTB
  and DBE became dominant in 2016 and were found to be the
  ``signature'' event of the GPU board failing resistor and a trigger
  for GPU replacement.}

@@

Initial processing updates an inventory by checking the Serial Number
(\rev{{\em SN}}) and Location of each GPU and creating or updating a
separate record for each contiguously observed
\rev{{\em{SN-Location}}} combination. Resulting GPU data records are
of the form shown in Fig.~\ref{fig:dataraw},
\begin{figure}[tb]
{\notsotiny
\begin{verbatim}
0323812007945 | c17-4c1s3n1 | 01/10/2013 02:54:45 | 02/02/2013 11:32:29
 | c13-1c1s3n3 | 01/21/2014 21:10:42 | 08/01/2017 02:15:02
 | c0-1c1s3n3 | 10/11/2013 15:57:33 | 10/12/2013 22:09:31
 | c21-1c2s5n0 | 03/19/2013 15:48:11 | 05/29/2013 11:54:11
0325216047736 | c18-4c1s5n1 | 03/07/2017 02:15:01 | 03/12/2019 03:19:11
0323812008856 | c5-4c0s7n0 | 01/10/2013 02:54:45 | 01/25/2013 15:29:58
 | c0-6c1s7n2 | 10/21/2013 14:28:19 | 10/28/2013 17:52:44
 | c3-3c1s5n0 | 05/29/2013 11:54:11 | 05/29/2013 11:54:11
 | c23-6c1s7n2 | 01/21/2014 21:10:42 | 11/02/2018 14:42:34
 | | DBE | 11/02/2018 14:42:34
\end{verbatim}
}
\caption{A few records of raw data produced from inventories \rev{and
    log files} that is processed further in our analysis.}
\label{fig:dataraw}
\end{figure}
where records for three are shown. Each record starts with a serial
number and locations are coded with {\tt c{\it col-row}c{\it
    cage}s{\it slot}n{\it node}}. The location references are cabinet
column (0-24), row (0-7), cage (0-2), slot (0-7), and node (0-3) with
respect to the layout shown in \rev{Fig.}~\ref{fig:layout}. \rev{Note
  that there is an extra empty column of cabinets between columns 10
  and 11 to accommodate a few ceiling support columns. To aid
  orientation in the figure, we mark an example location {\tt
    c17-4c1s3n1} with a yellow dot in each view of the figure.}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/TitanCabinet.png}
  \begin{minipage}{0.35\columnwidth}
    \includegraphics[width=\columnwidth]{figs/TitanBlade.png}
  \end{minipage}
  \hfil
  \begin{minipage}{0.49\columnwidth}
    \includegraphics[width=\columnwidth]{figs/TitanLayout.png}
  \end{minipage}
  \caption{\rev{Titan cabinet (front view and side view with air flow)
      and single blade layout (bottom left with {\em node} numbers)
      and floor layout (bottom right with {\em col} and {\em row}
      numbers). Cooling airflow is bottom to top within a cabinet,
      taking ambient air in mostly at the foot and relasing it back
      out at the top. Cooling fluid is passed through heat exchange
      evaporators, below Cage 0 and above Cage 2. Ambient air in the
      room is conditioned at nominal 75$^\circ$~F. The room air system
      has a dozen intakes near the ceiling and dumps cool air under
      the floor in several locations. A number of floor tiles are
      perforated.  Yellow dot marks GPU location {\tt c17-4c1s3n1}.}}
  \label{fig:layout}
\end{figure}

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Data Collection and Preprocessing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Collection and Preprocessing}
\label{section:dataprep}
Data on Titan GPU life times is constructed from two sources:
inventory runs and failure event \rev{log} records. Two types of
failure events were collected \rev{for this study}: Double Bit Error
(DBE) and Off the Bus (OTB). \rev{DBE is an error correcting code (ECC)
  detection in GPU memory, which can correct a single bit flip and detect
  but not correct a double bit flip. OTB is the loss of host CPU
  connection to the GPU. While other types of errors were recorded in
  log files, as is reported in \cite{7266836,tiwari15Understand}, OTB
  and DBE became dominant in 2016 and were found to be the
  ``signature'' event of the GPU board failing resistor and a trigger
  for GPU replacement.}

@@

Initial processing updates an inventory by checking the Serial Number
(\rev{{\em SN}}) and Location of each GPU and creating or updating a
separate record for each contiguously observed
\rev{{\em{SN-Location}}} combination. Resulting GPU data records are
of the form shown in Fig.~\ref{fig:dataraw},
\begin{figure}[tb]
{\notsotiny
\begin{verbatim}
0323812007945 | c17-4c1s3n1 | 01/10/2013 02:54:45 | 02/02/2013 11:32:29
 | c13-1c1s3n3 | 01/21/2014 21:10:42 | 08/01/2017 02:15:02
 | c0-1c1s3n3 | 10/11/2013 15:57:33 | 10/12/2013 22:09:31
 | c21-1c2s5n0 | 03/19/2013 15:48:11 | 05/29/2013 11:54:11
0325216047736 | c18-4c1s5n1 | 03/07/2017 02:15:01 | 03/12/2019 03:19:11
0323812008856 | c5-4c0s7n0 | 01/10/2013 02:54:45 | 01/25/2013 15:29:58
 | c0-6c1s7n2 | 10/21/2013 14:28:19 | 10/28/2013 17:52:44
 | c3-3c1s5n0 | 05/29/2013 11:54:11 | 05/29/2013 11:54:11
 | c23-6c1s7n2 | 01/21/2014 21:10:42 | 11/02/2018 14:42:34
 | | DBE | 11/02/2018 14:42:34
\end{verbatim}
}
\caption{A few records of raw data produced from inventories \rev{and
    log files} that is processed further in our analysis.}
\label{fig:dataraw}
\end{figure}
where records for three are shown. Each record starts with a serial
number and locations are coded with {\tt c{\it col-row}c{\it
    cage}s{\it slot}n{\it node}}. The location references are cabinet
column (0-24), row (0-7), cage (0-2), slot (0-7), and node (0-3) with
respect to the layout shown in \rev{Fig.}~\ref{fig:layout}. \rev{Note
  that there is an extra empty column of cabinets between columns 10
  and 11 to accommodate a few ceiling support columns. To aid
  orientation in the figure, we mark an example location {\tt
    c17-4c1s3n1} with a yellow dot in each view of the figure.}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/TitanCabinet.png}
  \begin{minipage}{0.35\columnwidth}
    \includegraphics[width=\columnwidth]{figs/TitanBlade.png}
  \end{minipage}
  \hfil
  \begin{minipage}{0.49\columnwidth}
    \includegraphics[width=\columnwidth]{figs/TitanLayout.png}
  \end{minipage}
  \caption{\rev{Titan cabinet (front view and left side view with air
      flow) and single blade layout (bottom left with {\em node}
      numbers) and floor layout (bottom right with {\em col} and {\em
        row} numbers). Cooling airflow is bottom to top within a
      cabinet, taking ambient air in mostly at the foot and relasing
      it back out at the top. Cooling fluid is passed through heat
      exchange evaporators, below Cage 0 and above Cage 2. Ambient air
      in the room is conditioned at nominal 75$^\circ$~F. The room air
      system has a dozen intakes near the ceiling and dumps cool air
      under the floor in several locations. A number of floor tiles
      are perforated.  Yellow dot marks GPU location {\tt
        c17-4c1s3n1}.}}
  \label{fig:layout}
\end{figure}

### 2020-06-05 discussion.tex
preparing for OLCF import

\rev{The failure rates of the old batch are not matched by the new
  batch nor by experience at other facilities with the same type of
  GPU components, making this an unexpected event that had
  considerable impact on operations and on availability of the
  system. On a positive but ironic note, the corrosion process in the
  failing resistors made the GPUs act as a sensitive instrument for
  obtaining long-term integrals of component heat dynamics and
  provided lessons learned. Here we dicsuss the conclusions we can
  draw from the experience on mitigation, which keeps the system
  operating at an acceptable level and possibly even restores some of
  the lost capability, and on long-term planning, which addresses such
  scenarios in future-generation supercomputers.}

>>>>

\rev{The failure rates of the old batch are not matched by the new
  batch nor by experience at other facilities with the same type of
  GPU components, making this an unexpected event that had
  considerable impact on operations and on availability of the
  system. On a positive but ironic note, the corrosion process in the
  failing resistors made the specific GPU batch act as sensitive
  instruments for obtaining cummulative trend information from
  component heat dynamics, giving the current analysis the strong
  signals observed, and provided lessons learned. Here we dicsuss the
  conclusions we can draw from the experience on mitigation, which
  keeps the system operating at an acceptable level and possibly even
  restores some of the lost capability, and on long-term planning,
  which addresses such scenarios in future-generation supercomputers.}

### 2020-06-05 survival.tex
preparing for OLCF import

\rev{An increasing average temperature gradient of about $4^\circ$C
  per cage is reported in \cite[Fig.10]{tiwari15Understand},
  independently confirming our airflow temperature
  interpretation. They also suggest that DBEs may be sensitive to
  temperature, although consider their evidence preliminary due to
  relatively low number of DBEs in 2014-1015 and high variability of
  $10^\circ$C to $15^\circ$C in their measured cage replicates.}

>>>>

\rev{An increasing average temperature gradient of about $4^\circ$C
  per cage is reported in \cite[Fig.10]{tiwari15Understand},
  independently confirming our airflow temperature
  interpretation. They also suggest that DBEs may be sensitive to
  temperature, although consider their evidence preliminary due to
  relatively low number of DBEs in 2014-2015 and high variability of
  $10^\circ$C to $15^\circ$C in their measured cage replicates.}
